Empirical Bayes analysis for detection of gene heterosis in RNAseq data

### 2014-12-17 ebayes_rnaseq_heterosis.tex
rewrote part of the ShrinkBayes paragraph and added comments

The {\tt ShrinkBayes} package implements fully Bayesian inference through integrated nested Laplace approximation (INLA) \citep{rue2009approximate,van2014shrinkbayes}. A limitation with INLA is that it only provides approximations to marginal posteriors for parameters or linear combinations of parameters rather than an approximation to the full posterior. Thus, we cannot directly estimate nonlinear quantities such as $P(\mu_{g3} < \min(\mu_{g1},\mu_{g2}))$. To circumvent this limitation, we reparameterized the independence model in Section \ref{s:model} following \cite{ji2014estimation}. Let $\phi_g = (\mu_{g1}+\mu_{g2})/2$ be the parental log-mean, $\alpha_g = (\mu_{g2}-\mu_{g1})/2$ be the half parental difference, and $\delta_g = \mu_{g3} - \phi_g$ be the hybrid effect. We estimated this model using {\tt ShrinkBayes} and constructed a statistic based on the posterior mean of these parameters. Specifically, the statistic is 0 if $|\hat{\delta}_f| \le |\hat{\alpha}_f|$, $P(\delta_f + \alpha_f > 0 |y)$ if $\hat{\delta}_f > -\hat{\alpha}_f \ge 0$, $P(\delta_f - \alpha_f > 0 |y)$ if $\hat{\delta}_f > \hat{\alpha}_f \ge 0$, $P(\delta_f - \alpha_f < 0 |y)$ if $\hat{\delta}_f < \hat{\alpha}_f \le 0$, and $P(\delta_f + \alpha_f < 0 |y)$ if $\hat{\delta}_f < -\hat{\alpha}_f \le 0$.

>>>>

The {\tt ShrinkBayes} package implements fully Bayesian inference through integrated nested Laplace approximation (INLA) \citep{rue2009approximate,van2014shrinkbayes}. A limitation of INLA is that it only provides approximations to marginal posteriors for parameters or linear combinations of parameters rather than an approximation to the full posterior. Thus, we cannot directly estimate nonlinear quantities such as $P(\mu_{g3} < \min(\mu_{g1},\mu_{g2}))$. To circumvent this limitation, we reparameterized the independence model in Section \ref{s:model} following \cite{ji2014estimation}.

% WML: your use of "following" in this last sentence is not clear to me.

% WML: Here, you use the phi-alpha-delta parameterization. You already defined phi and delta in previous sections. To avoid a conflict, you could easily change the notation here. It doesn't even have to be super intuitive because you only use it in this one paragraph.

 Let $\phi_g = (\mu_{g1}+\mu_{g2})/2$ be the parental log-mean, $\alpha_g = (\mu_{g2}-\mu_{g1})/2$ be the half parental difference, and $\delta_g = \mu_{g3} - \phi_g$ be the hybrid effect. After fitting the model with {\tt ShrinkBayes}, we constructed a pseudo-probability of heterosis for each gene. For gene $g$, we used the estimated posterior means $\hat{\delta}_g$ and $\hat{\alpha}_g$ of $\delta_g$ and $\alpha_g$, respectively, along with appropriate posterior probabilities $P(\cdot \mid y)$. Specifically, the pseudo-probability of heterosis for gene $g$ is zero if $|\hat{\delta}_f| \le |\hat{\alpha}_f|$, $P(\delta_f + \alpha_f > 0 |y)$ if $\hat{\delta}_f > -\hat{\alpha}_f \ge 0$, $P(\delta_f - \alpha_f > 0 |y)$ if $\hat{\delta}_f > \hat{\alpha}_f \ge 0$, $P(\delta_f - \alpha_f < 0 |y)$ if $\hat{\delta}_f < \hat{\alpha}_f \le 0$, and $P(\delta_f + \alpha_f < 0 |y)$ if $\hat{\delta}_f < -\hat{\alpha}_f \le 0$.
